---
title: "00_pf_sf_cohort_id"
author: "Chad H. Hochberg and Nicholas Ingraham"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Install Following Libraries For this Project
packages_to_install =
  c("tidyverse",
    "tidytable",
    "data.table",
    "collapse",
    "yaml",
    "fst",
    "arrow",
    "tableone",
    "viridis",
    "sandwich",
    "rprojroot", 
    "comorbidity")

#Can Load For Each Individual RMarkdown
packages_to_load =
  c(
    "data.table",
    "tidyverse",
    "collapse",
    "arrow",
    "data.table",
    "rprojroot")
options(dplyr.summarise.inform = FALSE)

fn_install_if_mi = function(p) {
  if (!requireNamespace(p, quietly = TRUE)) {
    try(install.packages(p, dependencies = TRUE), silent = TRUE)
  }
}

fn_load_quiet = function(p) {
  suppressPackageStartupMessages(library(p, character.only = TRUE))
}

invisible(lapply(packages_to_install, fn_install_if_mi))
invisible(lapply(packages_to_load,   fn_load_quiet))
options(dplyr.summarise.inform = FALSE)
rm(packages_to_install, packages_to_load, fn_install_if_mi, fn_load_quiet); gc()

#Set Working Director as the Project Root
project_root <- file.path( find_root(rprojroot::has_dir("CLIF_PFvsSF_Performance")),
  "CLIF_PFvsSF_Performance")
knitr::opts_knit$set(root.dir = project_root)
getwd()

#Use Dplyr functions as default
filter <- dplyr::filter
mutate <- dplyr::mutate
select <- dplyr::select
arrange <- dplyr::arrange
group_by <- dplyr::group_by
summarise <- dplyr::summarise
lead <- dplyr::lead
lag <- dplyr::lag
coalesce <- dplyr::coalesce
n <- dplyr::n
```


```{r Optimize Use of Cores - Code from Pat Lyons}
### threads and ram ------------------------------------------------------------

os_type   = Sys.info()[["sysname"]]
all_cores = parallel::detectCores(logical = TRUE)
all_cores = if (is.na(all_cores)) 1L else as.integer(all_cores)

get_ram_gb = function() {
  tryCatch({
    if (os_type == "Darwin") {
      # macOS: sysctl reports total RAM in bytes
      bytes = suppressWarnings(
        as.numeric(system("sysctl -n hw.memsize", intern = TRUE))
      )
      if (length(bytes) > 0 && !is.na(bytes)) bytes / 1024^3 else NA_real_
    } else {
      # Windows & Linux: ps is usually reliable
      mem_info <- ps::ps_system_memory()
      key <- intersect(c("available", "avail"), names(mem_info))
      val <- if (length(key) > 0) mem_info[[key]] / 1024^3 else NA_real_
  
      if (is.finite(val)) val else {
        # Linux fallback: read /proc/meminfo directly
        if (file.exists("/proc/meminfo")) {
          kb = suppressWarnings(
            as.numeric(system("awk '/MemAvailable/ {print $2}' /proc/meminfo", intern = TRUE))
          )
          if (length(kb) > 0 && !is.na(kb)) kb / 1024^2 else NA_real_
        } else {
          NA_real_
        }
      }
    }
  }, error = function(e) NA_real_)
}

avail_ram_gb = get_ram_gb()

## choose threads conservatively (00 is light; keep headroom) ------------------

reserve_cores   = 1L
gb_per_thread   = 0.50
max_by_cores    = max(1L, all_cores - reserve_cores)
max_by_memory   = if (is.finite(avail_ram_gb)) max(1L, floor(avail_ram_gb / gb_per_thread)) else max_by_cores
n_threads       = as.integer(max(1L, min(max_by_cores, max_by_memory, 8L)))
n_math_threads  = as.integer(max(1L, min(n_threads, 8L)))

## apply thread settings -------------------------------------------------------

data.table::setDTthreads(threads = n_threads)
collapse::set_collapse(nthreads  = n_threads)
options(arrow.use_threads        = TRUE)
Sys.setenv(ARROW_NUM_THREADS     = n_threads)
options(mc.cores                 = n_threads)

# concise summary
message(
  sprintf("Env OK | OS=%s | Cores=%d | Threads=%d | MathThreads=%d | Avail RAM≈%s GB",
          os_type, all_cores, n_threads, n_math_threads,
          ifelse(is.finite(avail_ram_gb), round(avail_ram_gb, 1), "NA"))
)
```
```{r Open Config File and Set Site Variables}
config = yaml::read_yaml(paste0(project_root, '/config/config.yaml'))
allowed_files = c("parquet", "csv", "fst")

### user entered site details ---------------------------------------------------
site             = tolower(config$site_lowercase)
file_type        = tolower(config$file_type)  
tables_location  = config$clif_data_path 
project_location = config$project_location

### file_type must be one of c(parquet, csv) -----------------------------------

if (!(file_type %in% allowed_files)) {
  stop(
    paste0(
      "Invalid '", file_type,
      "'. Expected one of: ", 
      paste(allowed_files, collapse = ", "), call. = FALSE
    )
  )
}

### pull time zone from site details -------------------------------------------

site_time_zone = config$site_timezone

check_timezone <- function(site_timezone, intended_timezone = Sys.timezone()) {
  if (site_timezone != intended_timezone) {
    cat("Warning: You specified", site_timezone, "but your system time zone is", intended_timezone, "\n")
    response <- readline(prompt = "Was that intentional? (yes/no): ")
    
    if (tolower(response) %in% c("no", "n")) {
      new_timezone <- readline(prompt = "Please enter the correct timezone: ")
      cat("Timezone updated to", new_timezone, "\n")
      return(new_timezone)
    } else {
      cat("Proceeding with the originally specified timezone:", site_timezone, "\n")
      return(site_timezone)
    }
    
  } else {
    cat("Timezone is set to", site_timezone, "\n")
    return(site_timezone)
  }
}

#Time Zone Check
result <- check_timezone(site_time_zone)

### file locations -------------------------------------------------------------

if (!dir.exists(paste0(project_location, "/proj_tables"))) {
  dir.create(paste0(project_location, "/proj_tables"))
}
if (!dir.exists(paste0(project_location, "/proj_output"))) {
  dir.create(paste0(project_location, "/proj_output"))
}
if (!dir.exists(paste0(project_location, "/proj_output/graphs"))) {
  dir.create(paste0(project_location, "/proj_output/graphs"))
}
```
```{r Required CLIF Table Paths}
#Vector of Required Tables
required_tables = 
  c(
    "patient", 
    "hospital_diagnosis",
    "hospitalization", 
    "adt", 
    "vitals", 
    "labs", 
    "medication_admin_continuous", 
    "respiratory_support",
    "code_status",
    "patient_assessments"
  )


clif_table_filenames = 
  list.files(
    path       = tables_location, 
    pattern    = paste0("^clif_.*\\.", file_type, "$"), 
    full.names = TRUE
  )

clif_table_basenames = 
  basename(clif_table_filenames) |>
  str_remove(paste0("\\.", file_type, "$")) |> # remove extension
  str_remove("^clif_") |>                      # remove leading 'clif_'
  str_remove("(_\\d{4}(_\\d{4})*)$")           # remove any date ranges in file name

table_file_map     = setNames(clif_table_filenames, clif_table_basenames)
missing_tables     = setdiff(required_tables, clif_table_basenames)
required_filenames = table_file_map[required_tables]

if (length(missing_tables) > 0) {
  stop(paste("Error: Missing required tables:", paste(missing_tables, collapse = ", ")))
} else {
  message("All required tables are present.")
}

rm(allowed_files, missing_tables); gc()

## load tables -----------------------------------------------------------------

if (file_type == "parquet") {
  data_list = lapply(required_filenames, open_dataset)
} else if (file_type == "csv") {
  data_list = lapply(required_filenames, \(f) read_csv_arrow(f))
} else if (file_type == "fst") {
  data_list = lapply(required_filenames, \(f) {
    tmp = read.fst(f, as.data.table = TRUE)
    arrow_table(tmp)
  })
} else {
  stop("Unsupported file format")
}

names(data_list) = names(required_filenames)
```
```{r Specify Global Variables - Time Period and Other Parameters}
STUDY_START <- as.POSIXct("2021-01-01", tz=site_time_zone) # Start time
STUDY_END <- as.POSIXct("2024-12-31", tz=site_time_zone) # End time

STABILITY_WINDOW_MINS <- 24 * 60 # Minimum number of minutes of NIV to be included
MAX_NIV_BREAK_MINS <- 2 * 60 # Maximum amount of minutes allowed off NIV within STABILITY_WINDOW_MINS from starting  
NIV_NAMES <- c("cpap", "nippv", "high flow nc") # Device names that are included as NIV

HOSPITAL_BLOCK_TIME_HRS <- 6 # Maximum number of hours allowed between hospitalizations to make them blocked together

# Age limits for study, defined by outlier_thresholds file
AGE_MIN <- 18
AGE_MAX <- 120

#SpO2, PaO2, FIO2 Outliers
SPO2_MIN <- 80
SPO2_MAX <- 100
PAO2_MIN <- 40
PAO2_MAX <- 700
FIO2_MIN <- 0.21
FIO2_MAX <- 1.00
PF_MAX <- 600

#VENT OUTLIERS
PEEP_MIN <- 0
PEEP_MAX <- 45
TV_MIN <- 50
TV_MAX <- 1500

#HIFNC OUTLIERS
HF_LPM_MIN <- 0
HF_LPM_MAX <- 90

#Respiratory Failure Duration (in Hours) - Minimum to Enter Study
RF_DURATION <- 24

#TIME from Spo2/Pao2 to Most Recent FiO2; 4 Hours
TIME_TO_SF_PF <- 4

#Blank Table to Build Consort Data
consort_table <- data.frame(
  step     = character(),
  included = numeric(),
  excluded = numeric(),   # should be numeric to match NA_real_
  stringsAsFactors = FALSE
)
```


#COHORT IDENTIFICATION
#Age >= 18
#Within Study Time Frame Noted Above
```{r Hospital Blocks Within Time Frame}
# Select only hospitalizations that fit the below criteria
# Only set the starting time admission (study end cutoff below as to not remove ends of hospital blocks)
hospitalizations <- data_list$hospitalization |>
   filter(
    age_at_admission >= AGE_MIN & age_at_admission < AGE_MAX,
    admission_dttm >= STUDY_START & admission_dttm <= STUDY_END,
    admission_dttm < discharge_dttm & !is.na(discharge_dttm)
  ) |>
  select(
    patient_id, hospitalization_id, admission_dttm, discharge_dttm,
    age_at_admission, discharge_category) |>
  arrange(patient_id, admission_dttm) |>
  distinct() |>
  compute()

# Collect as data table
hospitalizations <- as.data.table(hospitalizations)

# Order by patient ID and admission date/time
setorder(hospitalizations, patient_id, admission_dttm)

# Create column taht stores the time of previous discharge
hospitalizations[, prev_discharge := shift(discharge_dttm, n = 1), by = patient_id]

# Create column that stores the gap in hours between current admission and prior discharge
hospitalizations[, gap_hours := as.numeric(difftime(admission_dttm, prev_discharge, units = "hours"))]

# Flag a hospitalization if it represents a new hospital block
hospitalizations[, new_block := is.na(prev_discharge) | gap_hours > HOSPITAL_BLOCK_TIME_HRS]

# Compute cumulative block number per patient (e.g., first hospital block =1 , second = 2, etc.)
hospitalizations[, block_num := cumsum(replace_na(new_block, TRUE)), by = patient_id]

# Determine end of hospitalization time for discharge location
hospitalizations[, end_ts := coalesce(discharge_dttm, admission_dttm)]

#Remove helper columns
hospitalizations[, c("prev_discharge", "gap_hours", "new_block") := NULL]

# Create functions to return NA (if it is NA) or the min/max of a group
max_or_na <- function(x) if(all(is.na(x))) as.POSIXct(NA) else max(x, na.rm = TRUE)
min_or_na <- function(x) if(all(is.na(x))) as.POSIXct(NA) else min(x, na.rm = TRUE)

# Store information on grouped hospitalizations (hospital blocks)
hospitalizations <- hospitalizations[, `:=`(
  # Stores the patient age on first admission within block
  age_at_admission = min_or_na(age_at_admission),
  # Stores the start of the hospital block time
  final_admission_dttm = min_or_na(admission_dttm),
  # Stores the end of the hospital block time
  final_discharge_dttm = max_or_na(discharge_dttm),
  # Number of hospitalizations within a block
  num_hospitalizations = .N,
  # Stores the discharge location from last admission within block
  discharge_location = discharge_category[which.max(end_ts)]
), by = .(patient_id, block_num)]


# Create unique hospital block IDs
hospitalizations[, hospital_block_id := paste0(patient_id, "_", block_num)]

#Extract Unique IDs
hid_jid_crosswalk = select(hospitalizations, ends_with("id")) |> as_tibble()

# Create One Row per Hospitalization Block and Extract Info For Consort
hospitalizations <- hospitalizations[
  order(hospital_block_id, admission_dttm), 
  .SD[1], 
  by = hospital_block_id
]

#For Consort 
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "N Encounters for Adults In Registry During Study Time Frame", 
    included = length(unique(hospitalizations$hospital_block_id)), 
    excluded = NA_real_)
)
consort_table
```


#Admitted Via the ED
```{r Admitted Via the ED}
#Define If Patient First Admitted to ED
adt_info <- data_list$adt |>
  mutate(hospitalization_id=as.character(hospitalization_id)) |>
  inner_join(hid_jid_crosswalk) |> #Filter Out Hospitalizations Not initial Steps
  filter(tolower(location_category) %in% c('ed', 'ward', 'icu', 'stepdown')) |>
  arrange(hospital_block_id, in_dttm) |>
  group_by(hospital_block_id) |>
  as.data.table() #Faster To do Next Steps in Data Table

#Select Admission Hospital, Dishcarge Hospital and Type and Define if ED First
adt_info <- adt_info[, .(
  admit_hospital = hospital_id[1],
  discharge_hospital  = hospital_id[.N],
  admit_location_type = location_type[1],
  discharge_location_type  = location_type[.N],
  first_location_category = tolower(location_category[1]),
  ed_first            = tolower(location_category[1]) == "ed"
), by = hospital_block_id]

adt_info <- adt_info |>
  filter(ed_first==T)

#Merge with Hospitalizations - As Inner Join Starts Windowing Down the Cohort
hospitalizations <- hospitalizations |>
  collapse::join(adt_info, 
                 on = 'hospital_block_id',
                 how = 'inner')

#For Consort 
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "N Encounters Initially Presenting to ED", 
    included = length(unique(hospitalizations$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))
print(consort_table)

#Update hid_hid_crosswalk
hid_jid_crosswalk <- hid_jid_crosswalk |>
  join(hospitalizations, 
       on = 'hospital_block_id', 
       how = 'semi') #This will Filter to Hospital ID Blocks in Hospitalizations Table

rm(adt_info); gc()
```


#Define Respiratory Support
```{r Received HIFNC, NIV or IMV Within 7 Days of Admit for At Least 24 Hours}
#Load Respiratory Support and Filter to Relevant Time Frame 
#Perform Limited Respiratory Waterfill
#Assess For Periods of > 24 Hours of Respiratory Support

#To Asseess for Start of Respiratory Support within 7 Days of Admission Will COllect up to 8 days of Data (needs at least 24 hours)
day_cutoff <- 8*24*60*60 #Eight Days in Seconds 

resp_info <- data_list$respiratory_support |>
  #Just Evaluate Devices for Now
  filter(!is.na(device_category)) |>
  #Filter to Relevant Hospital Blocks
  inner_join(hid_jid_crosswalk %>% select(hospital_block_id, hospitalization_id)) |>
  #Filter to Relevant Time Frame (within 7 days) - THis means keep up to 8 days
  left_join(hospitalizations %>% 
              select(hospital_block_id, final_admission_dttm, final_discharge_dttm)) |>
  mutate(
    cutoff_dttm = final_admission_dttm + as.difftime(day_cutoff, units = "secs")
  ) |>
  filter(recorded_dttm <= cutoff_dttm) |>
  as.data.table() #Pull into R for Further Manipulation

#Filter to Only Encounters with Any IMV/NIV
resp_info <- resp_info[
  , .SD[any(tolower(device_category) %in% c("imv", "nippv", "cpap", "high flow nc", "high_flow_nc"))],
  by = hospital_block_id
]

#Address Some Outlier Issues Here
## Check fio2_set 
fio2_mean <- resp_info |> 
  select(fio2_set) |> 
  summarise(fio2_mean = mean(fio2_set, na.rm = TRUE)) |> 
  collect() |>
  pull() # fixing if its less than one # You will get a warning but it will be fixed on its own with IF statement 
if(fio2_mean > 1) { 
  cat('Mean FiO2 > 1, changing to fraction. fio2_mean=', fio2_mean, '\n')
  resp_info <- resp_info |> mutate(fio2_set = fio2_set / 100) |> compute() 
  } else {
  cat('Mean FiO2 <1, leaving as fraction. fio2_mean=', fio2_mean, '\n')
  }

#Address Initial Outliers
resp_info <- resp_info |>
  fmutate(
    # fio2_set
    fio2_set = fcase(
      fio2_set >= FIO2_MAX, NA_real_, 
      fio2_set <= FIO2_MIN, NA_real_,
      default = fio2_set),
    # Set tidal_volume_set
    tidal_volume_set = fcase(
      tidal_volume_set >= TV_MAX, NA_real_, 
      tidal_volume_set <= TV_MIN, NA_real_,
      default = tidal_volume_set),
    # PEEP SET
    peep_set = fcase(
      peep_set >= PEEP_MAX, NA_real_, 
      peep_set <= PEEP_MIN, NA_real_,
      default = peep_set),
    #For Nasal Cannula with high flow or high Fio2 mark as hiflow
    device_category=fifelse(
      device_category=='nasal_cannula' & lpm_set>=20 & fio2_set>=0.3, 
         'high flow nc', device_category))
    
#Waterfall Filling
resp_info <- resp_info |>
  tidytable::group_by(hospital_block_id, device_category) |>
  arrange(hospital_block_id, recorded_dttm) |>
  fmutate(
    fio2_carried=fio2_set,
    fio2_dttm=fifelse(!is.na(fio2_set), recorded_dttm, NA_POSIXct_),
    lpm_carried=lpm_set, 
    tidal_volume_carried=tidal_volume_set,
    peep_carried=peep_set
  ) |>
  tidytable::fill(fio2_carried, fio2_dttm, lpm_carried, tidal_volume_carried, peep_carried, .direction = 'downup') |>
  funique()

#Only Need to Evaluate Device Transitions and Last Row
resp_devices <- resp_info |>
  group_by(hospital_block_id) |>
  mutate(device_change=fcase(
    row_number()==1, 1L, #First Row Counts as Device Transition ('New Device')
    device_category!=lag(device_category), 1L, #Otherwise if Device Prior is Different It is a new Device
    default=0L), 
    first_row=fifelse(row_number()==1, 1L, 0L),
    last_row=fifelse(row_number()==n(), 1L, 0L)
  ) 

#Now Filter
resp_devices <- resp_devices |>
  filter((device_change==T | is.na(device_change)) | first_row==1L | last_row==1L)

#Now Define Time in Respiratory Failure; This Means at Least 24 Continuous Hours of IMV, NIPPV or HFNC (or combination)
resp_episodes <- resp_devices |>
  fmutate(
    #Respiratory Support For Purposes of this Study
    resp_support =fcase(
      tolower(device_category) %in% c('imv', 'nippv', 'cpap'), 1L,
      tolower(device_category) %in% c('high flow nc', 'high_flow_nc') & lpm_carried>=30, 1L,
      default = 0L)) 

#Repeat Process to Filter Just to Changes
resp_episodes <- resp_episodes |>
  group_by(hospital_block_id) |>
  mutate(
    changing_support = fcase(
      row_number()==1 & resp_support==1L, 1L,
      lag(resp_support)!=resp_support, 1L,
      default=0L),
    first_row=fifelse(row_number()==1, 1L, 0L),
    last_row=fifelse(row_number()==n(), 1L, 0L)
  ) 

#Now Filter to Only On/OFF Transitions in Respiratory Support
resp_episodes <- resp_episodes |>
  filter((changing_support==T | is.na(changing_support)) | first_row==1L | last_row==1L) |>
  group_by(hospital_block_id) |>
  mutate(
    next_dttm = fifelse(
      last_row==T, 
      final_discharge_dttm,
      lead(recorded_dttm)),
   #Number Episodes Based on New Starts THis will be For Cumulative Duration of Resp Support Calculation
   new_start=fcase(
      (first_row==1L & resp_support==1L), 1L, 
      (changing_support==T & resp_support==1L), 1L,
      default=0L),
    resp_support_episode_num=cumsum(new_start==1L),
    resp_support_episode_num=fifelse(resp_support==0L, 0L, resp_support_episode_num),
    resp_support_duration = fifelse(
      resp_support==1L, 
      as.numeric(difftime(next_dttm, recorded_dttm, units='hours')), 
      NA_real_)) |>
    #Now Group by HOspital Block and Episode Number for Total Duration
    group_by(hospital_block_id, resp_support_episode_num) |>
    fmutate(resp_support_duration=sum(resp_support_duration, na.rm=T)) |>
    ungroup()
  
#Select First Episode That is >=24 Hours and Start Before Cutoff Day
resp_episodes_on_study <- resp_episodes |>
  fmutate(
    resp_failure_start=fcase(
       resp_support==1L, recorded_dttm),
    resp_failure_stop=resp_failure_start+dhours(resp_support_duration)) |>
  filter(resp_support==1L & resp_support_duration>=RF_DURATION,
         resp_failure_start<=cutoff_dttm) |> #Have to Do this Again As we Used Data up to Day 8
  group_by(hospital_block_id) |>
  arrange(hospital_block_id, recorded_dttm) |>
  ffirst() |>
  ungroup() |>
  fselect(hospital_block_id, resp_support_episode_num, resp_support_duration, resp_failure_start, resp_failure_stop)

#Update Consort Table
#For Consort 
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "N Encounters Receiving More Than 24 Continuous Hours of Respiratory Support in First 7 Days", 
    included = length(unique(resp_episodes_on_study$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))
print(consort_table)

#Filter Resp Info To Just These Patients
resp_info <- resp_info |>
  semi_join(resp_episodes_on_study, by='hospital_block_id')

#Filter hid_jid_cosswalk to just these patients
hid_jid_crosswalk <- hid_jid_crosswalk |>
  semi_join(resp_episodes_on_study, by='hospital_block_id')

#REmove Tables no Longer Neededd
rm(resp_devices); gc()
```


Exclusions
#First Airway is Tracheostomy
#No Available Data on FiO2 in the First 24 Hours
#Not Comfort Care Status in First 24 Hours of Respiratory Support
#No Eligible PF or SF Measures (that is PF or SF measured within 4 Hours of FIO2 Documentation)

```{r Exclusions}
#Add Tracheostomy Present in First 24 Hours as Variable
#Indicate if No FiO2 Data Available
trach_fio2 <- resp_info |>
  left_join(resp_episodes_on_study, by='hospital_block_id') |>
  as.data.table() |>
  filter(
      recorded_dttm>=resp_failure_start & 
        recorded_dttm<=resp_failure_start+dhours(24)) |>
  group_by(hospital_block_id) |>
  fmutate(tracheostomy = fifelse(is.na(tracheostomy), 0L, tracheostomy)) |> 
  fmutate(
    trach_first_24=fifelse(mean(tracheostomy, na.rm=T)>0, 1L, 0L),
    no_fio2=(sum(!is.na(fio2_set))==0)
  ) |>
  slice_head(n=1) |>
  select(hospital_block_id, trach_first_24, no_fio2) |>
  ungroup()

#Now Apply to Respiratory Episodes Table
resp_episodes_on_study <- resp_episodes_on_study |>
  semi_join(trach_fio2 |> filter(no_fio2==F), by='hospital_block_id')
#Track for Consert
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "n, excluded for no FiO2 Recorded in First 24 Hours", 
    included = length(unique(resp_episodes_on_study$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))

resp_episodes_on_study <- resp_episodes_on_study |>
  semi_join(trach_fio2 |> filter(trach_first_24==F), by='hospital_block_id')
#Track for Consert
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "n, excluded for no Tracheostomy Present in First 24 Hours", 
    included = length(unique(resp_episodes_on_study$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))
#Adjust Current Hospitalizations in Cohort
hid_jid_crosswalk <- hid_jid_crosswalk |>
  semi_join(resp_episodes_on_study, by='hospital_block_id')
rm(trach_fio2)

#Not Comfort Care Status in First 24 Hours of Respiratory Support
dni_equivalent <- c(
  'and',
  'dnr/dni',
  'dnar/dni'
)
codes = 
  data_list[["code_status"]] |>
  semi_join(hid_jid_crosswalk, by='patient_id') |>
  left_join(hid_jid_crosswalk, by='patient_id', multiple=T) |>
  select(hospital_block_id, start_dttm, code_status_category) |>
  collect() |>
  funique()

codes = 
  join(resp_episodes_on_study, codes, how = "left", on='hospital_block_id', multiple = T) |>
  #Code Status Within 7 Days Prior to and Up to 24 Hours After Respiratory Failure Start
  fsubset(start_dttm >= resp_failure_start - lubridate::ddays(7)) |>
  fsubset(start_dttm <= resp_failure_start + lubridate::dhours(24)) |>
  roworder(hospital_block_id, start_dttm) |>
  fgroup_by(hospital_block_id) |>
  fsummarize(code_status = flast(code_status_category)) |>
  #If no code status present Within 7 Days Prior to and 24 Hours After Respiratory Failure Start Prseume Full
  fmutate(limited_code_status_24h = if_else(tolower(code_status) %in% dni_equivalent, 1L, 0L, 0L))

#Join Back wiht Respiratory Episodes
resp_episodes_on_study <- resp_episodes_on_study |>
  left_join(codes, by = 'hospital_block_id') |>
  fmutate(limited_code_status_24h=fifelse(is.na(limited_code_status_24h), 0L, limited_code_status_24h)) |>
  filter(limited_code_status_24h==0L)
  
#Track for Consert
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "n, excluded for DNI Status In First 24 Hours of Resp Failure", 
    included = length(unique(resp_episodes_on_study$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))
#Adjust Current Hospitalizations in Cohort
hid_jid_crosswalk <- hid_jid_crosswalk |>
  semi_join(resp_episodes_on_study, by='hospital_block_id')
rm(codes)

```

#Exposures:
Primary: Worst P/F and Worst S/F in First 24 Hours
Secondary: Median P/F and Median S/F in First 24 Hours
Groupings: For Worst PF/SF Measures Resp Support Mode is Mode on Which PF/SF Was Measured
For Median PF/SF it is the most support HIFNC>NIPPV>IMV in First 24 Hours

#At End of this Block We Will Have Data to Implement Final Exclusion of No Eligible PF or SF Measures in First 24 Hours
```{r Generate PF and SF Exposurem Measures}
#Table for Respiratory Info Filtered to Patients in Cohort
resp_first_24 <- resp_info |>
  semi_join(hid_jid_crosswalk, by = 'hospital_block_id') |>
  left_join(resp_episodes_on_study, by='hospital_block_id') |>
  fsubset(recorded_dttm>=resp_failure_start & recorded_dttm<=resp_failure_start+dhours(24)) |>
  fgroup_by(hospital_block_id) |>
  fmutate(
    resp_support_rank=fcase(
      tolower(device_category) %in% c('high flow nc', 'high_flow_nc'), 0L,
      tolower(device_category) %in% c('nippv', 'cpap'), 1L,
      tolower(device_category)=='imv', 2L,
      default = 0L),
    resp_support_rank=max(resp_support_rank),
    max_resp_support24=fcase(
      resp_support_rank==0L, 'high_flow_nc',
      resp_support_rank==1L, 'nippv', 
      resp_support_rank==2L, 'imv'
    )) |>
  ungroup()

#Bring in ABG Data
abgs = data_list[["labs"]] |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  semi_join(hid_jid_crosswalk, by='hospitalization_id') |>
  filter(tolower(lab_category) %in% c(
    'ph_arterial',
    'po2_arterial',
    'pco2_arterial')) |>
  left_join(hid_jid_crosswalk, by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study, by='hospital_block_id', multiple=T) |>
  select(hospital_block_id, lab_category, lab_value_numeric, lab_collect_dttm, resp_failure_start) |>
  filter(lab_collect_dttm>=resp_failure_start & lab_collect_dttm<=resp_failure_start+dhours(24)) |>
  as.data.table() |>
  funique() |>
  tidytable::pivot_wider(names_from = lab_category, values_from=lab_value_numeric, 
                         id_cols = c('hospital_block_id', 'lab_collect_dttm'), 
                         values_fn = median) |>
  #Outlier Filtering
  mutate(po2_arterial=fifelse(po2_arterial>PAO2_MAX | po2_arterial<PAO2_MIN, NA_real_, po2_arterial)) |>
  rename(recorded_dttm=lab_collect_dttm) |>
  select(hospital_block_id, recorded_dttm, ph_arterial, po2_arterial, pco2_arterial)

#Bring in SpO2 Data
spo2 = data_list[["vitals"]] |>
    mutate(hospitalization_id = as.character(hospitalization_id)) |>
  semi_join(hid_jid_crosswalk, by='hospitalization_id') |>
  filter(tolower(vital_category)=='spo2') |>
  left_join(hid_jid_crosswalk, by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study, by='hospital_block_id', multiple=T) |>
  filter(
    recorded_dttm>=resp_failure_start & recorded_dttm<=resp_failure_start+dhours(24),
    vital_value>=SPO2_MIN & vital_value<=SPO2_MAX) |>
  as.data.table() |>
  funique() |>
  rename(spo2=vital_value) |>
  select(hospital_block_id, recorded_dttm, spo2)
  
#Merge Into Respiratory Table
resp_first_24 <- resp_first_24 |>
  tidytable::full_join(abgs, by = c('hospital_block_id', 'recorded_dttm')) |>
  tidytable::full_join(spo2, by = c('hospital_block_id', 'recorded_dttm')) |>
  group_by(hospital_block_id) |>
  fill(device_category, .direction = 'downup') |>
  group_by(hospital_block_id, device_category) |>
  mutate(
    fio2_carried=fio2_set, 
    fio2_dttm=fifelse(!is.na(fio2_set), fio2_dttm, NA_POSIXct_)) |>
  arrange(hospital_block_id, recorded_dttm) |>
  fill(fio2_carried, fio2_dttm, .direction = 'downup') |>
  #Only Need to Keep Pao2 and SpO2 Measures
  ungroup()

#Mark if Measure (Spo2 or Pao2 Measured within 4 Hours of Fio2 Documentation)
resp_first_24_long <- resp_first_24 |>
  #Can Use a PF or SF if Measured Within 4 Hours of Last Fio2 Documentation
  #Can Use SF if SpO2 <=97%
  mutate(
    eligible_measure_by_time=fifelse(
         as.numeric(difftime(recorded_dttm, fio2_dttm, units = "hours"))<=TIME_TO_SF_PF, 1L, 0L, 0L),
    eligible_sf_by_spo2=fifelse(
      spo2<=97, 1L, 0L, 0L),
    eligible_spo2=fifelse(eligible_measure_by_time==1L & eligible_sf_by_spo2==1L, 1L, 0L, 0L)) |>
  #Now Define Worst S/F and Worst P/F in the First 24 Hours
  group_by(hospital_block_id, eligible_measure_by_time) |>
  mutate(
    pf_ratio=po2_arterial/fio2_carried,
    pf_ratio=fifelse(pf_ratio>=PF_MAX, NA_real_, pf_ratio),
    pf_24_min=fifelse(eligible_measure_by_time==1, min(pf_ratio, na.rm=T, default = NA_real_), NA_real_),
    pf_24_mean=fifelse(eligible_measure_by_time==1, mean(pf_ratio, na.rm=T, default = NA_real_), NA_real_),
    pf_24_median=fifelse(eligible_measure_by_time==1, median(pf_ratio, na.rm=T, default = NA_real_), NA_real_),
    #Mark Which PF Is Minimum
    this_pf_temp=fifelse(pf_ratio==pf_24_min, 1L, 0L, 0L)
  ) |>
  group_by(hospital_block_id) |>
  mutate(
    across(
      c(pf_24_min, pf_24_mean, pf_24_median),~ ifelse(is.finite(.), ., NA_real_)),
    this_pf = row_number() == min(which(this_pf_temp == 1), na.rm = TRUE), 
    min_pf_dttm=fifelse(this_pf==1, recorded_dttm, NA_POSIXct_), 
    min_pf_device=fifelse(this_pf==1, device_category, NA_character_),
    ph_pf=fifelse(this_pf==1, ph_arterial, NA_real_),
    paco2_pf=fifelse(this_pf==1, pco2_arterial, NA_real_),
    min_pf_pao2=fifelse(this_pf==1, po2_arterial, NA_real_),
    min_pf_fio2=fifelse(this_pf==1, fio2_carried, NA_real_)
    ) |>
  fill(max_resp_support24, min_pf_dttm, min_pf_device, pf_24_min, pf_24_mean, pf_24_median, ph_pf, paco2_pf, min_pf_pao2, min_pf_fio2, .direction='downup') |>
  #Now Spo2
  group_by(hospital_block_id, eligible_spo2) |>
  mutate(
    sf_ratio=spo2/fio2_carried, 
    sf_24_min=fifelse(eligible_spo2==1, min(sf_ratio, na.rm=T, default = NA_real_), NA_real_),
    sf_24_mean=fifelse(eligible_spo2==1, mean(sf_ratio, na.rm=T, default = NA_real_), NA_real_),
    sf_24_median=fifelse(eligible_spo2==1, median(sf_ratio, na.rm=T, default = NA_real_), NA_real_),
    #Mark Which SF Is Minimum
    this_sf_temp=fifelse(sf_ratio==sf_24_min, 1L, 0L, 0L)) |>
  group_by(hospital_block_id) |>
  mutate(
    across(
      c(sf_24_min, sf_24_mean, sf_24_median),~ ifelse(is.finite(.), ., NA_real_)),
    this_sf = row_number() == min(which(this_sf_temp == 1), na.rm = TRUE), 
    min_sf_dttm=fifelse(this_sf==1, recorded_dttm, NA_POSIXct_), 
    min_sf_device=fifelse(this_sf==1, device_category, NA_character_),
    min_sf_spo2=fifelse(this_sf==1, spo2, NA_real_),
    min_sf_fio2=fifelse(this_sf==1, fio2_carried, NA_real_)
    ) |>
  fill(min_sf_dttm, min_sf_device, sf_24_min, sf_24_mean, sf_24_median, min_sf_spo2, min_sf_fio2, .direction='downup') |>
  ungroup() |>
  select(hospital_block_id, recorded_dttm, eligible_measure_by_time, eligible_spo2, max_resp_support24, fio2_dttm, fio2_carried, fio2_set, spo2, po2_arterial, min_pf_dttm, min_pf_device, pf_24_min, min_pf_pao2, min_pf_fio2, pf_24_mean, pf_24_median, ph_pf, paco2_pf, this_pf, po2_arterial, pf_ratio, ph_arterial, pco2_arterial,spo2, this_sf, min_sf_dttm, min_sf_device, sf_24_min, sf_24_mean, sf_24_median, min_sf_spo2, min_sf_fio2)

rm(abgs, spo2, resp_first_24); gc()
```

```{r Exclude Those with no Eligible PF/SF}
#Exclude if NO Eligible PF or SF (Within 4 Hours of FIO2 Documentation)
exclude <- resp_first_24_long |>
  group_by(hospital_block_id) |>
  filter(sum(eligible_measure_by_time, na.rm=T)==0) |>
  ungroup() |>
  distinct(hospital_block_id)

if (nrow(exclude)>0) {
  hid_jid_crosswalk <- hid_jid_crosswalk |> filter(!hospital_block_id %in% exclude$hospital_block_id)
  resp_first_24_long <- resp_first_24_long |>
    filter(!hospital_block_id %in% exclude$hospital_block_id)
} else {
  cat('\nNo Patients Excluded for Lack of SF or PF Eligible Measurements')
}

#For Consort 
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "n, with Eligible PF and/or SF Ratios", 
    included = length(unique(hid_jid_crosswalk$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))

rm(exclude)
```
```{r Finally Randomly Select One Encounter For Each Patient}
set.seed(32284)
randomly_select <- hid_jid_crosswalk |>
  distinct(patient_id, hospital_block_id) |>
  group_by(patient_id) |>
  slice_sample(n=1)

hid_jid_crosswalk <- hid_jid_crosswalk |>
  filter(hospital_block_id %in% randomly_select$hospital_block_id)

#For Consort 
consort_table <- rbind(
  consort_table,
  data.frame(
    step = "n, Randomly Selected 1 Encounter Per Patient", 
    included = length(unique(hid_jid_crosswalk$hospital_block_id)), 
    excluded = NA_real_
)) |>
  #This Step Keeps Track of Exclusions
  mutate(excluded=case_when(
    row_number()==n() ~ lag(included)-included, 
    TRUE ~ excluded
  ))

#Filter respiratory Tables to the Final Cohort
resp_episodes_on_study <- resp_episodes_on_study |>
  filter(hospital_block_id %in% hid_jid_crosswalk$hospital_block_id)
resp_first_24_long <- resp_first_24_long |>
  filter(hospital_block_id %in% hid_jid_crosswalk$hospital_block_id)

rm(randomly_select); gc()
```

```{r Create Summary of PF/SF Measures and Exposure Categories}
#Summary Table of PF and SF Measures at Patient Level 
resp_first_24_summary <- resp_first_24_long |>
  group_by(hospital_block_id) |>
   mutate(
    n_spo2=sum(!is.na(spo2)),
    n_fio2=sum(!is.na(fio2_set)), 
    n_pao2=sum(!is.na(po2_arterial)),
    n_eligible_sf_spo2_lt98 = max(cumsum(fifelse(eligible_spo2 == 1, 1, 0, 0)), na.rm = TRUE), 
    n_eligible_sf_spo2_gt97 = max(cumsum(fifelse(eligible_measure_by_time == 1 & spo2 > 97, 1, 0, 0)), na.rm = TRUE),
    n_eligible_pf = max(cumsum(fifelse(eligible_measure_by_time == 1 & !is.na(po2_arterial), 1, 0, 0)), na.rm = TRUE)
  ) |>
  filter(row_number()==1) |>
  ungroup() |>
  select(
    hospital_block_id, max_resp_support24, min_pf_dttm, min_pf_device, pf_24_min, min_pf_pao2, min_pf_fio2, pf_24_mean, pf_24_median, 
    ph_pf, paco2_pf, min_sf_dttm, min_sf_device, sf_24_min, sf_24_mean, sf_24_median, min_sf_spo2, min_sf_fio2, n_spo2, n_fio2, n_pao2,     n_eligible_sf_spo2_lt98, n_eligible_sf_spo2_gt97, n_eligible_pf)

#Define Respiratory Support at Time of Inclusion
resp_first_24_summary <- resp_first_24_summary |>
  mutate(
    resp_support_category=tolower(fcase(
      !is.na(pf_24_min), min_pf_device,
      !is.na(sf_24_min), min_sf_device,
      default = max_resp_support24
  )),
    resp_support_category=str_replace_all(resp_support_category, " ", "_"), #Snake Case
    exposure_group=fcase(
      !is.na(pf_24_min) & is.na(sf_24_min), 'pf_only',
      !is.na(pf_24_min) & !is.na(sf_24_min), 'sf_pf',
      is.na(pf_24_min) & !is.na(sf_24_min), 'sf_only',
      is.na(pf_24_min) & is.na(sf_24_min), 'undefined_oxygenation'
    ))
```


```{r Covariates - Demographics and Hospital Outcomes}
demos_outcomes <- hospitalizations |>
  semi_join(hid_jid_crosswalk, by='hospitalization_id') |>
  select(patient_id, hospital_block_id, final_admission_dttm, final_discharge_dttm, age_at_admission, discharge_category, discharge_location, admit_hospital_id=admit_hospital, discharge_hospital_id=discharge_hospital) |>
  #Define Hospital Outcomes
  mutate(
    inhosp_death=fifelse(tolower(discharge_category)=='expired', 1L, 0L),
    hospice_discharge=fifelse(tolower(discharge_category)=='hospice', 1L, 0L),
    inhosp_death_hospice=fifelse(inhosp_death==1 | hospice_discharge==1, 1L, 0L))


patient = data_list[['patient']] |>
  semi_join(demos_outcomes, by='patient_id') |>
  select(patient_id, race_category, ethnicity_category, sex_category) |>
  mutate(
    female = if_else(tolower(sex_category) == 'female', 1L, 0L),
    new_race_cat = case_when(
      tolower(race_category) == 'asian' ~ 'asian', 
      tolower(race_category) == 'black or african american' ~ 'black',
      tolower(race_category) == 'white' & tolower(ethnicity_category) %in% c('non-hispanic', 'unknown') ~ 'white',
      tolower(race_category) == 'white' & tolower(ethnicity_category) == 'hispanic' ~ 'hispanic',
      is.na(race_category) & tolower(ethnicity_category) == 'hispanic' ~ 'hispanic',
      TRUE ~ 'other'
    )
  ) |>
  as.data.table()
  
cohort_build <- demos_outcomes |>
  left_join(patient, by='patient_id')

rm(hospitalizations, patient, demos_outcomes)
```

```{r Covariates - Elixhauser}
## add elixhauser -------------------------------------------------------------- code from Patrick Lyons

### vector common codes to avoid for RAM's sake (to repl w/ keep list) ---------

unused_vect = c(
  "Z79.899",
  "E78.5",
  "Z87.891",
  "K21.9",
  "Z20.822",
  "F17.210",
  "Z79.01",
  "N17.9"
)

### only relevant codes from relevant encounters -------------------------------

elix = 
  dplyr::filter(data_list$hospital_diagnosis, hospitalization_id %in% hid_jid_crosswalk$hospitalization_id) |>
  dplyr::filter(!toupper(diagnosis_code) %in% unused_vect) |>
  # Join hospital_block_id early
  left_join(hid_jid_crosswalk %>% select(hospitalization_id, hospital_block_id), by = 'hospitalization_id') |>
  dplyr::select(hospital_block_id, poa_present, diagnosis_code) |>  # Use hospital_block_id
  dplyr::collect() |>
  funique()


### assign elixhauser diagnosis dummies ----------------------------------------

elix = 
  comorbidity::comorbidity(
    elix, 
    id      = "hospital_block_id",  # Use hospital_block_id as the ID
    code    = "diagnosis_code", 
    map     = "elixhauser_icd10_quan", 
    assign0 = T
  )

############# TOOK THIS OUT... DID AGGREGATION TO ENCOUNTERS ABOVE
# ### link to joined-hosp-id -----------------------------------------------------
# 
# elix = elix |>
#   left_join(hid_jid_crosswalk %>% select(hospitalization_id, hospital_block_id), by = 'hospitalization_id') |>
#   fgroup_by(hospital_block_id) 
#   fmax()

### calculate vw scores --------------------------------------------------------

vw_elixhauser       = comorbidity::score(elix, weights = "vw", assign0 = T)
elixhauser_count =  elix |>
   mutate(elixhauser_count = rowSums(across(2:32), na.rm = TRUE)) |>
   fselect(hospital_block_id, elixhauser_count)
elix      = cbind(elix, vw_elixhauser = vw_elixhauser) |> fselect(hospital_block_id, vw_elixhauser) |>
  left_join(elixhauser_count %>% select(hospital_block_id, elixhauser_count), by = 'hospital_block_id')



cohort_build <- cohort_build |>
  left_join(elix, by='hospital_block_id')

rm(elix, elixhauser_count, vw_elixhauser); gc()
```

```{r Covariates - BMI and SOFA Score}
#Calculate BMI
bmi <- data_list[['vitals']] |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  filter(tolower(vital_category) %in% c('weight_kg', 'height_cm')) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  #Filter Outliers
  filter((tolower(vital_category)=='height_cm' & vital_value>=76.2 & vital_value<=244) |
         (tolower(vital_category)=='weight_kg' & vital_value>=20 & vital_value<=1100)) |>
  as.data.table() |>
  #Pick Height and Weight Closest to Enrollment
  #First Calculate Time Difference
  mutate(
    time_diff=as.duration(as.POSIXct(resp_failure_start)-as.POSIXct(recorded_dttm))/dhours(1),
  #Define if Before or After t_entrollment
    before_enrollment=fifelse(time_diff>=0, 1, 0)) |>
  group_by(patient_id, hospital_block_id, vital_category) |> #With this Grouping You keep Closest Before and After resp_failure_start
  mutate(keep=fifelse(
    abs(time_diff)==min(abs(time_diff)), 1, 0
  )) |>
  filter(keep==1) |>
  #Now Select Height/Weight (Preferrably This is Prior to Enrollment)
  group_by(patient_id, hospital_block_id) |>
  mutate(study_height_cm=fcase(
    keep==1 & before_enrollment==1 & vital_category=='height_cm', vital_value
  )) |>
  mutate(study_weight_kg=fcase(
      keep==1 & before_enrollment==1 & vital_category=='weight_kg', vital_value
  )) |>
  fill(study_height_cm, study_weight_kg, .direction = 'downup') |>
  #IF missing Can Use First Value After Enrollment
  mutate(study_height_cm=fifelse(
    is.na(study_height_cm) & keep==1 & 
      before_enrollment==0 & vital_category=='height_cm', 
    vital_value, study_height_cm
  )) |>
  mutate(study_weight_kg=fifelse(
     is.na(study_weight_kg) &  keep==1 & before_enrollment==0 & vital_category=='weight_kg', 
     vital_value, study_weight_kg
  )) |>
  fill(study_height_cm, study_weight_kg, .direction = 'downup') |>
  filter(row_number()==1) |>
  ungroup() |>
  #Calculate BMI
  mutate(bmi=study_weight_kg/((study_height_cm/100)^2)) |>
  select(patient_id, hospital_block_id, study_height_cm, study_weight_kg, bmi)

#Calculate Norepinephrine Equivalent
pressor_data <- data_list[['medication_admin_continuous']] |>
  filter(med_group=='vasoactives') |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(admin_dttm>=resp_failure_start & admin_dttm<=(resp_failure_start+dhours(24))) |>
  left_join(bmi %>% select(hospital_block_id, study_weight_kg), by ='hospital_block_id') |> #For Weights
  arrange(patient_id, hospital_block_id, admin_dttm) |>
  as.data.table() |>
  mutate(med_dose=as.character(med_dose),
        med_dose = as.numeric(parse_number(med_dose)), 
  #Define if a Pressor Needs Weight Based Dosing
        med_dose_unit=tolower(med_dose_unit),
        needs_wt_based=fifelse((med_dose_unit=="mcg/min" | med_dose_unit=="mg/min" | med_dose_unit=="ng/min"), 1, 0)) |>
  #Filter if Epinephrine or Phenylephrine is Given in Bolus Dsing 'mg' as this is code dosing
  filter(!(med_category=='epinephrine' & med_dose_unit %in% c('mg', 'mcg')), 
         !(med_category=='phenylephrine' & med_dose_unit=='mcg'),
         !(med_dose_unit %in% c('mg', 'ml/hr', 'units', 'milli-units/kg/hr'))) |>
  mutate(wt_dose=fcase(
    needs_wt_based==0, med_dose,
    needs_wt_based==1, med_dose/study_weight_kg)) |>
  #Filter Out Outliers
  filter(!(med_category=='norepinephrine' & wt_dose>=6),
         !(med_category=='epinephrine' & wt_dose>=6),
         !(med_category=='vasopressin' & wt_dose>=0.2)) |>
  group_by(hospital_block_id) |>
  mutate(norepi_dose=fcase(med_category=="norepinephrine" , wt_dose)) |>
  fill(norepi_dose, .direction ="down") |>
  mutate(norepi_on=fcase(
    norepi_dose > 0 , 1,
    (is.na(norepi_dose) | norepi_dose==0) , 0 )) |>
  mutate(norepi_dose=fifelse(is.na(norepi_dose), 0, norepi_dose)) |> #Need to Have Dose be Zero if No Dose Got Carried Forward
  mutate(epi_dose=fcase(
    med_category=="epinephrine" , wt_dose
  )) |>
  fill(epi_dose, .direction = "down") |>
  mutate(epi_on=fcase(
    epi_dose>0 , 1,
    (is.na(epi_dose) | epi_dose==0) , 0 )) |>
  mutate(epi_dose=fifelse(is.na(epi_dose), 0, epi_dose)) |>
  mutate(phenyl_dose=fcase(
    med_category=="phenylephrine" , wt_dose
  )) |>
  fill(phenyl_dose, .direction = "down") |>
  mutate(phenyl_on=fcase(
    phenyl_dose>0 , 1,
    (is.na(phenyl_dose) | phenyl_dose==0) , 0 )) |>
  mutate(phenyl_dose=fifelse(is.na(phenyl_dose), 0, phenyl_dose)) |>
  mutate(vaso_dose=fcase(
    med_category=="vasopressin" , wt_dose
  )) |>
  fill(vaso_dose, .direction = "down") |>
  mutate(vaso_on=fcase(
    vaso_dose>0 , 1,
    (is.na(vaso_dose) | vaso_dose==0) , 0 )) |>
  mutate(vaso_dose=fifelse(is.na(vaso_dose), 0, vaso_dose)) |>
  mutate(angiotensin_dose=fcase(
    med_category=="angiotensin" , wt_dose
  )) |>
  fill(angiotensin_dose, .direction = "down") |>
  mutate(angiotensin_on=fcase(
    angiotensin_dose>0 , 1,
    (is.na(angiotensin_dose) | angiotensin_dose==0) , 0 )) |>
  mutate(angiotensin_dose=fifelse(is.na(angiotensin_dose), 0, angiotensin_dose)) |>
  mutate(dopa_dose=fcase(
    med_category=="dopamine" , wt_dose
  )) |>
  fill(dopa_dose, .direction = "down") |>
  mutate(dopa_on=fcase(
    dopa_dose>0 , 1,
    (is.na(dopa_dose) | dopa_dose==0) , 0 )) |>
  mutate(dopa_dose=fifelse(is.na(dopa_dose), 0, dopa_dose)) |>
  mutate(dobut_dose=fcase(
    med_category=="dobutamine" , wt_dose
  )) |>
  fill(dobut_dose, .direction = "down") |>
  mutate(dobut_on=fcase(
    dobut_dose>0 , 1,
    (is.na(dobut_dose) | dobut_dose==0) , 0 )) |>
  mutate(milrinone_dose=fcase(
    med_category=="milrinone" , wt_dose
  )) |>
  fill(milrinone_dose, .direction = "down") |>
  mutate(milrinone_on=fcase(
    milrinone_dose>0 , 1,
    (is.na(milrinone_dose) | milrinone_dose==0) , 0 )) |>
  mutate(isoproterenol_dose=fcase(
    med_category=="isoproterenol" , wt_dose
  )) |>
  fill(isoproterenol_dose, .direction = "down") |>
  mutate(isoproterenol_on=fcase(
    isoproterenol_dose>0 , 1,
    (is.na(isoproterenol_dose) |isoproterenol_dose==0) , 0 )) |>
  ungroup() |>
  #See Kotani et al for Norepi dose equivalents for Epi, Phenyl, Vaso, Dopa, Angiotensin)
  mutate(pressor_on=(norepi_on + epi_on + phenyl_on + vaso_on + dopa_on + angiotensin_on), 
         norepi_equivalent=norepi_dose + epi_dose+ vaso_dose*2.5+ angiotensin_dose*0.0025+ dopa_dose/100+ phenyl_dose*0.06,
         sofa_pressor=case_when(dopa_dose>15 | epi_dose>0.1 | norepi_dose>0.1 ~4, 
                                dopa_dose>5 | (epi_dose>0 & epi_dose <=0.1) | (norepi_dose>0 & norepi_dose<=0.1) ~3, 
                                (dopa_dose>0 & dopa_dose<=5) | dobut_dose>0~2 , 
                                pressor_on==0~0), 
         sofa_pressor=case_when(is.na(sofa_pressor)~0, is.na(sofa_pressor)==F ~sofa_pressor)) 

#Select max sofa pressor in 24h period
sofa_pressors <- pressor_data |>
  group_by(hospital_block_id) |>
  mutate(max_sofa_pressor=max(sofa_pressor, na.rm = TRUE)) |>
  filter(row_number()==1) |>
  select(hospital_block_id, max_sofa_pressor) |> #, norepi_dose, epi_dose, vaso_dose, angiotensin_dose, dopa_dose, phenyl_dose) |>
  ungroup()

#Now Select Max NEE In this 24 Hour Period 
norepi_equivalent <- pressor_data |>
  group_by(hospital_block_id) |>
  mutate(max_norepi_equivalent=max(norepi_equivalent, na.rm = TRUE)) |>
  filter(row_number()==1) |>
  mutate(nee_pressor_dose=case_when(
    max_norepi_equivalent==0~0, 
    max_norepi_equivalent>0 & max_norepi_equivalent<=0.1~1, 
    max_norepi_equivalent>.1~2)) |>
  select(hospital_block_id, nee_pressor_dose, max_norepi_equivalent) |> #, norepi_dose, epi_dose, vaso_dose, angiotensin_dose, dopa_dose, phenyl_dose) |>
  ungroup()

# Calculate lowest GCS for sofa in 24h of Resp Failure Assessment
sofa_gcs <- data_list[['patient_assessments']] |>
  filter(tolower(assessment_category) %in% c("gcs_total")) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(recorded_dttm>=resp_failure_start & recorded_dttm<=(resp_failure_start+dhours(24))) |>
  #Filter Outliers  
  filter(numerical_value<=15 & numerical_value>=3) |>
  mutate(sofa_gcs=case_when(numerical_value==15~0,  
                            numerical_value==13 | numerical_value==14~1, 
                            numerical_value>=10 & numerical_value<=12~2, numerical_value>=6 & numerical_value<=9~3,
                            numerical_value<6~4)) |> 
  group_by(hospital_block_id) |>
  mutate(max_sofa_gcs=max(sofa_gcs, na.rm = TRUE)) |>
  select(hospital_block_id, max_sofa_gcs) |> 
  as.data.table() |>
  group_by(hospital_block_id) |>
  filter(row_number()==1) |>
  ungroup()

# Calculate lowest RASS for sofa in 24h prior to enrollment
sofa_rass <- data_list[['patient_assessments']] |>
  filter(tolower(assessment_category) %in% c("rass")) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(recorded_dttm>=resp_failure_start & recorded_dttm<=(resp_failure_start+dhours(24))) |>
  #Filter Outliers  
  filter(numerical_value<=4 & numerical_value>-6) |>
  mutate(max_sofa_rass=
             case_when(numerical_value>-1~0,  
                       numerical_value==-1~1, 
                       numerical_value==-2~2, 
                       numerical_value==-3~3, 
                       numerical_value==-4 | numerical_value==-5~4)) |> 
  as.data.table() |>
  select(hospital_block_id, max_sofa_rass) |>
  group_by(hospital_block_id) |>
  fmax()

#Calculate Lowest MAP 24h prior to enrollment for SOFA
sofa_map <-  data_list[['vitals']] |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  filter(tolower(vital_category) %in% c("map")) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(recorded_dttm>=resp_failure_start & recorded_dttm<=(resp_failure_start+dhours(24))) |>
  #Filter Outliers MAP>250
  filter(vital_value<=250) |>
  mutate(max_sofa_map=case_when(
    vital_value<70~1, 
    vital_value>=70~0)) |>
  select(hospital_block_id, max_sofa_map) |>
  as.data.table() |>
  group_by(hospital_block_id) |>
  fmax()

#For Labs Allow For Carried Forward Labs Up to 72 Hours
#Calculate highest bilirubin in 24h prior to enrollment for sofa
sofa_bili <-  data_list[['labs']] |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  filter(lab_category %in% c('bilirubin_total')) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(lab_collect_dttm>=(resp_failure_start-dhours(72)) & lab_collect_dttm<=(resp_failure_start+dhours(24))) |>
  #Filter Outliers  
  filter(lab_value_numeric<=80) |>
  mutate(
    bili_24=if_else(
      lab_collect_dttm>=resp_failure_start & lab_collect_dttm<=(resp_failure_start+dhours(24)), 1L, 0L),
    sofa_bili=case_when(
        lab_value_numeric<1.2 ~ 0,  
        lab_value_numeric>=1.2 & lab_value_numeric<2 ~ 1, 
        lab_value_numeric>=2 & lab_value_numeric<6 ~ 2, 
        lab_value_numeric>=6 & lab_value_numeric<12 ~ 3, 
        lab_value_numeric>=12 ~ 4)) |> 
  select(hospital_block_id, lab_collect_dttm, sofa_bili, bili_24) |>
  as.data.table() |>
  group_by(hospital_block_id, bili_24) |>
  mutate(
    max_sofa_bili=if_else(
      bili_24==1L, fmax(sofa_bili), NA_real_)) |>
  group_by(hospital_block_id) |>
  arrange(hospital_block_id, lab_collect_dttm) |>
  fill(max_sofa_bili, .direction='downup') |>
  mutate(max_sofa_bili=fifelse(
      is.na(max_sofa_bili) & row_number()==n(), sofa_bili, max_sofa_bili)) |>
  fill(max_sofa_bili, .direction='downup') |>
  filter(row_number()==1) |>
  select(hospital_block_id, max_sofa_bili) |> 
  ungroup()

#Calculate lowest platelets in 24h prior to enrollment for sofa
sofa_plt <-  data_list[['labs']] |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  filter(lab_category %in% c('platelet_count')) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(lab_collect_dttm>=(resp_failure_start-dhours(72)) & lab_collect_dttm<=(resp_failure_start+dhours(24))) |>
  #Filter Outliers  
  filter(lab_value_numeric<=2000) |>
  mutate(
    plt_24=if_else(
      lab_collect_dttm>=resp_failure_start & lab_collect_dttm<=(resp_failure_start+dhours(24)), 1L, 0L),
    sofa_plt=case_when(
        lab_value_numeric<20~4,  
        lab_value_numeric>=20 & lab_value_numeric<50~3, 
        lab_value_numeric>=50 & lab_value_numeric<100~2, 
        lab_value_numeric>=100 & lab_value_numeric<150~1, 
        lab_value_numeric>=150~ 0)) |> 
  select(hospital_block_id, lab_collect_dttm, sofa_plt, plt_24) |>
  as.data.table() |>
  group_by(hospital_block_id, plt_24) |>
  mutate(
    max_sofa_plt=if_else(
      plt_24==1L, fmax(sofa_plt), NA_real_)) |>
  group_by(hospital_block_id) |>
  arrange(hospital_block_id, lab_collect_dttm) |>
  fill(max_sofa_plt, .direction='downup') |>
  mutate(max_sofa_plt=fifelse(
      is.na(max_sofa_plt) & row_number()==n(), sofa_plt, max_sofa_plt)) |>
  fill(max_sofa_plt, .direction='downup') |>
  filter(row_number()==1) |>
  select(hospital_block_id, max_sofa_plt) |> 
  ungroup()


#Calculate highest cr in 24h prior to enrollment for sofa
sofa_cr <-  data_list[['labs']] |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  filter(lab_category %in% c('creatinine')) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start, resp_failure_stop)) |>
  filter(lab_collect_dttm>=(resp_failure_start-dhours(72)) & lab_collect_dttm<=(resp_failure_start+dhours(24))) |>
  #Filter Outliers  
  filter(lab_value_numeric<=20) |>
  mutate(
    cr_24=if_else(
      lab_collect_dttm>=resp_failure_start & lab_collect_dttm<=(resp_failure_start+dhours(24)), 1L, 0L),
    sofa_cr=case_when(
      lab_value_numeric<1.2 ~ 0,  
      lab_value_numeric>=1.2 & lab_value_numeric<2 ~ 1, 
      lab_value_numeric>=2 & lab_value_numeric<3.5 ~ 2, 
      lab_value_numeric>=3.5 & lab_value_numeric<5 ~ 3, 
      lab_value_numeric>=5 ~ 4)) |> 
  select(hospital_block_id, lab_collect_dttm, sofa_cr, cr_24) |>
  as.data.table() |>
  group_by(hospital_block_id, cr_24) |>
  mutate(
    max_sofa_cr=if_else(
      cr_24==1L, fmax(sofa_cr), NA_real_)) |>
  group_by(hospital_block_id) |>
  arrange(hospital_block_id, lab_collect_dttm) |>
  fill(max_sofa_cr, .direction='downup') |>
  mutate(max_sofa_cr=fifelse(
      is.na(max_sofa_cr) & row_number()==n(), sofa_cr, max_sofa_cr)) |>
  fill(max_sofa_cr, .direction='downup') |>
  filter(row_number()==1) |>
  select(hospital_block_id, max_sofa_cr) |> 
  ungroup()

#Merge BMI data into Cohort Build
cohort_build <- cohort_build |>
  left_join(bmi, by=c('patient_id', 'hospital_block_id')) |>
  left_join(norepi_equivalent, by=c('hospital_block_id')) |>
  #Will interpret missing information on vasopressors as evidence of no vasopressors (no vasoactives listed in MAR)
  fmutate(
    nee_pressor_dose=fifelse(
      is.na(nee_pressor_dose), 0, nee_pressor_dose),
    max_norepi_equivalent=fifelse(
      is.na(max_norepi_equivalent), 0, max_norepi_equivalent))

#Process Cohort SOFA scores and keep track of SOFA missingness prior to imputation as 0
cohort_sofa <- resp_first_24_summary |>
  mutate(
    max_sofa_resp=fcase(
        # Use actual PaO2/FiO2 if available; Otherwise SF
        !is.na(pf_24_min) & pf_24_min < 100 & 
          max_resp_support24 %in% c('high_flow_nc', 'imv','nippv'), 4,
        !is.na(pf_24_min) & pf_24_min < 200 & 
          max_resp_support24 %in% c('high_flow_nc', 'imv','nippv'), 3,
        !is.na(pf_24_min) & pf_24_min < 300, 2,
        !is.na(pf_24_min) & pf_24_min < 400, 1,
        !is.na(pf_24_min) & pf_24_min >= 400, 1,
        # Use S/F Equivalents if PF Not Available, Use CUtoffs from Matthay Global ARDS
        !is.na(sf_24_min) & sf_24_min < 148 & 
          max_resp_support24 %in% c('high_flow_nc', 'imv','nippv'), 4,
        !is.na(sf_24_min) & sf_24_min < 235 & 
          max_resp_support24 %in% c('high_flow_nc', 'imv','nippv'), 3,
        !is.na(sf_24_min) & sf_24_min < 315, 2,
        !is.na(sf_24_min) & sf_24_min < 400, 1,
        !is.na(sf_24_min) & sf_24_min >= 400, 1,
        #IF both S/F and P/F Missing Use 1 for NIV, Hiflow, 2 for Vent
        is.na(sf_24_min) & is.na(pf_24_min) & 
          max_resp_support24 %in% c('imv'), 2,
        is.na(sf_24_min) & is.na(pf_24_min) & 
          max_resp_support24 %in% c('high_flow_nc', 'nippv'), 1,
        default=NA_real_)) |>
  select(hospital_block_id, max_sofa_resp) |>
  left_join(sofa_bili, by=c('hospital_block_id')) |>
  left_join(sofa_cr, by=c('hospital_block_id')) |>
  left_join(sofa_plt, by=c('hospital_block_id')) |>
  left_join(sofa_pressors, by=c('hospital_block_id')) |>
  left_join(sofa_map, by=c('hospital_block_id')) |>
  left_join(sofa_gcs, by=c('hospital_block_id')) |>
  left_join(sofa_rass, by=c('hospital_block_id')) |>
  #GENERAT SOFA CV AND SOFA CNS SCORES
  mutate(
    max_sofa_cv=fcase(
      max_sofa_pressor>=2, max_sofa_pressor, 
      default =max_sofa_map), 
   max_sofa_cns=fifelse(
     !is.na(max_sofa_gcs), max_sofa_gcs, max_sofa_rass))
    

#Describe SOFA Missingness and Save CSV of SOFA Missingness Prior to Imputation as 0
sofa_missing <- cohort_sofa %>%
  pivot_longer(
    cols = -hospital_block_id, # Retain hospital_block_id as ID
    names_to = "variable",      # Name of the column to contain the names of the variables
    values_to = "value"         # Name of the column to contain the values of the variables
  ) |>
  group_by(variable) |>
  summarise(
    missing=sum(is.na(value)),
    missing_fraction=missing/n()
  )
#Write to CSV
write_csv(sofa_missing, paste0(project_location, '/proj_output/', site, '_sofa_missing.csv'))

#Now Calculate Max Score
cohort_sofa <- cohort_sofa |>
  mutate(max_sofa_total=
           rowSums(across(c(max_sofa_resp, max_sofa_bili, max_sofa_cr, 
                            max_sofa_plt, max_sofa_cv, max_sofa_cns)), na.rm = TRUE))

cohort_build <- cohort_build |>
  left_join(cohort_sofa)

rm(bmi, cohort_sofa, sofa_missing, sofa_bili, sofa_cr, sofa_gcs, sofa_map, sofa_plt, sofa_pressors, sofa_rass, pressor_data, norepi_equivalent) ; gc()

```

```{r Covaraites - Hospital ID and ICU or IMC}
#Create Data Frame For Defining This
location_data <- data_list$adt |>
  select(hospitalization_id, hospital_id, hospital_type, in_dttm, out_dttm, location_category, location_type) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk %>% select(hospitalization_id, hospital_block_id),  by = 'hospitalization_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start)) |>
  group_by(hospital_block_id) |>
  arrange(hospital_block_id, in_dttm) |>
  as.data.table()

#Hospital First - This will the Hospital Where the Patient Was at the Start of Respiratory Failure
location_data <- location_data |>
  group_by(hospital_block_id, hospital_id) |>
  mutate(
    last_hospital_time=max(out_dttm), 
    hospital_id=fifelse(
      resp_failure_start<last_hospital_time, hospital_id, NA_character_),
    hospital_type=fifelse(
      resp_failure_start<last_hospital_time, hospital_type, NA_character_)) |>
  group_by(hospital_block_id) |>
  arrange(hospital_block_id, in_dttm) |>
  fill(hospital_id, hospital_type, .direction = 'downup') |>
  #Patient in ICU Within 48 Hours of Respiratory Failure Start?
  mutate(
    keep=fcase(
      in_dttm<=resp_failure_start & out_dttm>=resp_failure_start, 1L,
      in_dttm>=resp_failure_start & in_dttm<=(resp_failure_start+dhours(48)), 1L, 
      resp_failure_start<min(in_dttm), 1L, #Few Patients have Initiation of Vitals Before 'in_dttm' listed (generally by minutes)
      default=0L),
    keep=cumsum(keep)) |>
  #Only Need to Consider Locations within 488 Hours Respiratory Start
  filter(
    keep>=1L,
    in_dttm<=(resp_failure_start+dhours(48))) |>
  mutate(
    imc_admit = as.integer(any(tolower(location_category) == 'stepdown')),
    icu_admit = as.integer(any(tolower(location_category) == 'icu')),
    ward_admit = as.integer(imc_admit == 0L & icu_admit == 0L & any(tolower(location_category) == 'ward')),
    location_type = if_else(location_type=='NA', NA_character_, location_type), #Hopkins Patch
    icu_type = coalesce(location_type)  # Keep location_type if not missing, else NA
  ) |>
  select(hospital_block_id, hospital_id, hospital_type, icu_admit, imc_admit, ward_admit, icu_type) |>
  fill(icu_type, .direction = 'updown') |>
  filter(row_number()==1) |>
  ungroup()
  
check_missing_adt <- data_list$adt |>
  select(hospitalization_id, hospital_id, hospital_type, in_dttm, out_dttm, location_category, location_type) |>
  semi_join(hid_jid_crosswalk,  by = 'hospitalization_id') |>
  left_join(hid_jid_crosswalk %>% select(hospitalization_id, hospital_block_id),  by = 'hospitalization_id') |>
  anti_join(location_data, by = 'hospital_block_id') |>
  left_join(resp_episodes_on_study %>% select(hospital_block_id, resp_failure_start)) |>
  as.data.table()

#Bring Into Cohort Build
cohort_build <- cohort_build |>
  left_join(location_data, by='hospital_block_id')

rm(location_data); gc()
```

```{r Complete Analytic Cohort}
analytic_cohort <- cohort_build |>
  left_join(resp_episodes_on_study, by='hospital_block_id') |>
  left_join(resp_first_24_summary, by='hospital_block_id')
```

```{r Save Analytic Datasets}
write_parquet(analytic_cohort, paste0(project_location, '/proj_tables/analytic_cohort.parquet'))
write_parquet(resp_first_24_long, paste0(project_location, '/proj_tables/resp_first_24_long.parquet'))
write_csv(consort_table, paste0(project_location, '/proj_output/',site,'_consort_data.csv'))
```

